---
title: "String Manipulation & Patterns with `stringr`"
output: html_document
---

# Introduction: What is `stringr`?

`stringr` is an R package from the `tidyverse` family designed for string manipulation. It simplifies working with text data by providing consistent, user-friendly functions for tasks like detecting patterns, counting characters, replacing text, and using regular expressions ("regex") to identify complex patterns.

## Why `stringr` is Important for Data Science

In biological research, string manipulation is frequently required to:

  - Analyze DNA and protein sequences (e.g., counting specific amino acids or nucleotide motifs)
  - Extract information from patient records (e.g., identifying family history patterns)
  - Standardize data (e.g., harmonizing labels for health statuses)

The consistent syntax of `stringr` functions (`stringr::str_*`) makes it intuitive for beginners and powerful enough for more advanced string operations.

# Getting Started

Make sure you have the `tidyverse` package installed. `stringr` is included within `tidyverse`, so loading the `tidyverse` will load `stringr` automatically.

```{r setup, include = FALSE}
# install.packages("tidyverse")
library(tidyverse)
```

## Load Example Data

We'll work with two datasets:

  - Patient History (`patient_hx`): This fictional dataset includes family history data.
  - Proteins (`proteins`): Contains amino acid sequences for various proteins.

```{r}
# Load patient data
patient_hx_url <- "https://raw.githubusercontent.com/tidybiology/tidybiology.org/main/data/phx.csv"
patient_hx <- readr::read_csv(patient_hx_url)

# Load proteins data
# install.packages("remotes")
# remotes::install_github("tidybiology/tidybiology", force = TRUE)

library(tidybiology)
data(proteins)
```

# Exercises

## Exercise 1: Counting Substrings in Protein Data

The goal here is to count occurrences of specific patterns in strings. This skill is essential when analyzing text data in biology, such as counting amino acids in protein sequences.

Given the `proteins` dataset, count the occurrences of lysine (represented by "K") in each sequence, and return the protein with the highest lysine content.

  - Use the `stringr::str_count()` function
  - Specify the pattern "K" to count each occurrence of lysine

```{r exercise1-answer_chunk, purl=FALSE}
# Count occurrences of lysine (K) in each sequence
proteins

# Identify protein with the highest lysine content

```

## Exercise 2: Detecting Signatures in Protein Sequences

Learn to detect the presence of specific signatures in protein sequences using `stringr::str_detect()`. This skill is essential for identifying functional regions within proteins, such as binding sites or structural domains.

Given the `proteins` dataset, identify which sequences contain the "KR" pattern (lysine and arginine), commonly associated with post-translational modification sites. Add a logical column (TRUE/FALSE) `has_kr` that marks sequences containing this pattern. Finally, count the number of columns with this pattern.

```{r exercise2-answer_chunk, purl=FALSE}
# Detect "KR" pattern
proteins

# Count proteins with "KR" pattern (has_kr == TRUE)

```

## Exercise 3: Replacing Text to Standardize Data

Practice replacing text within strings using `stringr::str_replace()`. Data often needs standardizing, such as consolidating health status terms for consistent analysis.

### Difference `str_replace()` and `str_replace_all()` 

The functions `str_replace()` and `str_replace_all()` both replace patterns within strings, but they differ in how many replacements they make:

  - `str_replace()`: Replaces only the first instance of the pattern in each string.
  - `str_replace_all()`: Replaces all instances of the pattern in each string.

Example: 

```{r}
example_sequences <- c("MKRTYYKK", "GGKTRK", "KPGGRK")

# Replace only the first instance of "K" in each sequence
stringr::str_replace(example_sequences, "K", "X")

# Replace all instances of "K" in each sequence
stringr::str_replace_all(example_sequences, "K", "X")
```

Using the `patient_hx` dataset, standardize the values in the `health_status` column by replacing "Normal blood sugar and cholesterol" with "Normal". This reduces complexity in downstream analysis.

```{r exercise3-answer_chunk, purl=FALSE}
# Standardize "Normal blood sugar and cholesterol" to "Normal"
patient_hx

```

## Basics of Regular Expressions

Regular expressions (regex) are sequences of characters that define search patterns, primarily for string matching and manipulation. Here’s a rundown of regex basics, organized by categories that cover essential concepts and symbols.

### Literal Characters

The simplest form of regex. A pattern like `"cat"` will match the exact string "cat" anywhere it appears.

### Metacharacters

These are special characters with unique meanings. To use them as regular characters, they must be escaped with a backslash (`\`).

Common Metacharacters:

`.` (dot): Matches any single character (e.g., `c.t` matches "cat", "cut", etc.)
`^`: Anchors the pattern to the beginning of a string (e.g., `^cat` matches "catfish" but not "concatenate")
`$`: Anchors the pattern to the end of a string (e.g., `fish$` matches "catfish" but not "fishing")

### Character Classes

Character classes let you specify a set of characters that you want to match.

Examples:

`[abc]`: Matches any one of the characters "a", "b", or "c"
`[a-z]`: Matches any lowercase letter from "a" to "z"
`[^abc]`: Matches any character except "a", "b", or "c"

### Quantifiers

Quantifiers specify how many times a character or group should appear.

Examples:

`*`: Matches 0 or more of the preceding character (e.g., `ca*t` matches "ct", "cat", "caat", etc.)
`+`: Matches 1 or more of the preceding character (e.g., `ca+t` matches "cat", "caat", but not "ct")
`?`: Matches 0 or 1 of the preceding character (e.g., `ca?t` matches "cat" or "ct")
`{n}`: Matches exactly n occurrences of the preceding character (e.g., `a{3}` matches "aaa")
`{n,m}`: Matches between n and m occurrences of the preceding character (e.g., `a{2,4}` matches "aa", "aaa", or "aaaa")

### Groups and Alternation

Parentheses `()`: Used to group parts of a regex to apply quantifiers or to extract specific parts of matches.
`|` (Alternation): Works like an OR operator (e.g., `cat|dog` matches "cat" or "dog").

Example:

The pattern `(cat|dog)s?` matches "cat", "cats", "dog", or "dogs".

### Escaping Special Characters

To match metacharacters literally, escape them with a backslash (`\`).

Example:

To match `.` as a literal dot, use `\.`

### Anchors

Anchors don’t match characters but instead position within the string.

`^`: Start of the string.
`$`: End of the string.

Example:

The pattern `^The` matches strings starting with "The".

### Advanced Techniques (Lookaheads and Lookbehinds)

Lookaheads and lookbehinds allow you to match a pattern only if it's followed or preceded by another pattern, without including that pattern in the match.

Example:

Positive Lookahead `(?=...)`: Matches if the specified pattern is followed by another pattern (e.g., `apple(?= pie)` matches "apple pie" and "green apple pie" but not "apple tart" or "apple").
Negative Lookahead `(?!...)`: Matches if the specified pattern is not followed by another pattern (e.g., `\\d+(?=%)` matches "25%" and "100%" but not "50 dollars" or "just 10")

## Exercise 4: Complex Pattern Matching with Regular Expressions

Understand the basics of regular expressions for complex pattern matching using `stringr::str_count()`. This is useful for finding patterns across various cases, such as identifying specific family members in patient data.

Using the `patient_hx` dataset, identify how many patients have a father in their family history but not a grandfather. Combine regex with `str_count()` to see the frequency of matches.

```{r exercise4-answer_chunk, purl=FALSE}
# Count occurrences of "Father" or "father" at the start, excluding "grandfather"
family_history

```

## Exercise 5: Filtering Based on Detected Patterns

Learn to filter data based on detected patterns, crucial for narrowing down datasets, like filtering proteins with a specific signature.

Filter the `proteins` dataset to only include proteins with the "CANNTG" signature (E-box motif), **where "N" can be any amino acid**. Create a new filtered dataframe called `e_box_proteins`.

Use `str_detect()` within `dplyr::filter()` to isolate specific rows based on pattern matches.

```{r exercise5-answer_chunk, purl=FALSE}
# Filter proteins with the E-box signature "CANNTG"
proteins

```

# Advanced Regex

Regular expressions ("regex") offer a way to match patterns in text by combining characters and symbols with special meanings. Regex is extremely versatile in biological data analysis, especially in complex sequence analyses and patient data extraction tasks.

Below, we cover some advanced regex techniques, including `lookaheads` and `lookbehinds`, character `classes`, and `quantifiers`. These exercises will build upon earlier string manipulation skills and apply regex for more complex biological pattern-matching tasks.

## Advanced Exercise 1: Positive and Negative Lookaheads

Learn to use lookaheads to match patterns based on characters that follow (or don’t follow) a specified point. This is particularly useful when needing to locate a sequence with specific contexts.

Find DNA sequences where "TATA" is followed by an A or T, without including "TATA" that is followed by other nucleotides.

```{r}
# Sample DNA sequences
dna_sequences <- data.frame(
  sequence = c("TATACGGA", "TATATGCA", "GGTATAGG")
)

# Detect "TATA" followed by A or T using positive lookahead

```

## Advanced Exercise 2: Negative Lookaheads

Use negative lookaheads to match patterns based on characters that do not follow a specified point, helpful for excluding unwanted patterns.

Identify patients with "Mother" in their health history but exclude cases mentioning "Grandmother".

```{r}
# Sample family history data
family_history <- data.frame(
  relationship = c("Mother", "grandmother", "Grandmother", "Motherly")
)

# Match "Mother" but exclude "Grandmother" using negative lookahead

```

## Advanced Exercise 3: Character Classes and Quantifiers

Use character classes and quantifiers for matching multiple characters or sequences of specific lengths.

Identify proteins containing any three consecutive polar amino acids (e.g., S, T, Y, N, Q). This example applies character classes and quantifiers for biological sequence patterns.

```{r}
# Sample protein sequences
protein_sequences <- data.frame(
  sequence = c("ASTYNGKK", "GGYSTYYT", "KPGGRKYY")
)

# Match sequences with three consecutive polar amino acids

```

# Resources for Further Learning

[Stringr Documentation](https://stringr.tidyverse.org/)

