---
title: "Dplyr Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
library(tidyverse)
```

## Required package and data set

Run the code in the code chunk below to load the necessary data set - `cytomegalovirus`
```{r}
#install.packages("medicaldata")
library(medicaldata)

# For more information on this package and the various data sets it contains, visit -
# https://github.com/higgi13425/medicaldata
# this is a treasure trove of interesting data sets!

data("cytomegalovirus") # loads this data set 
View(cytomegalovirus) # View this data set
?cytomegalovirus # launches the help page for this data set
```

## Please answer the following questions 

Spend ~15-20 minutes working through this section in pairs, and we will then go through the questions as a group. 

Take a look at the `cytomegalovirus` data frame. How many different types of variables does it contain? 

Hint: Take a glimpse at your data  

What are the dimensions of this data frame?  

```{r}
glimpse(cytomegalovirus)
```

How would you select just the columns `time.to.transplant`, `time.to.agvhd`, `time.to.cgvhd`, and `time.to.cmv` from the `cytomegalovirus` data frame?

**Bonus**: Can you identify a helpful function that will help you do this with less typing? Go through this list - https://dplyr.tidyverse.org/reference/dplyr_tidy_select.html

```{r}
cytomegalovirus %>% 
  dplyr::select(time.to.transplant, time.to.agvhd, time.to.cgvhd, time.to.cmv)
```

How would you select everything between (and including) `CD34.dose` and `cmv`?

```{r}
cytomegalovirus %>% 
  dplyr::select(CD34.dose:cmv)
```

Filter the `cytomegalovirus` data frame so that it only contains information for patients with acute myeloid leukemia

```{r}
cytomegalovirus %>% 
  dplyr::filter(diagnosis == "acute myeloid leukemia")
```

Now use `filter()` to only keep data for female patients with acute myeloid leukemia 

First try this using multiple filter statements, and then try without writing multiple filter statements 

```{r}
# with multiple filter statements
cytomegalovirus %>%
  dplyr::filter(diagnosis == "acute myeloid leukemia") %>% 
  dplyr::filter(sex == 1)
  
# one statement
cytomegalovirus %>%
  dplyr::filter(diagnosis == "acute myeloid leukemia",
                sex == 1)
```

Filter out information for patients who have a **below average** time.to.transplant

Hint: Use the base R function `mean()` for this question  

**WARNING**: Is your output an empty data frame? You will need to change the value of one of the arguments of the `mean()` function. Go through the function's documentation to decide which one  

```{r}
cytomegalovirus %>% 
  dplyr::filter(time.to.transplant > mean(time.to.transplant, na.rm = TRUE))
```

Create a new variable called `time.to.transplant.years` that contains the time.to.transplant for each patient in *years*

As a bonus, re-order the resulting data frame in descending order of `time.to.transplant.years` 

```{r}
cytomegalovirus %>% 
  dplyr::mutate(time.to.transplant.years = time.to.transplant/12) %>% 
  dplyr::arrange(desc(time.to.transplant.years))
```

Let's now find out which diseases have the longest time to transplant. We'll do this by working out the average `time.to.transplant` for each diagnosis

```{r}
cytomegalovirus %>% 
  dplyr::group_by(diagnosis) %>% 
  dplyr::summarise(mean(time.to.transplant, na.rm = TRUE))
```

Modify the code above to add a column that has the counts for each diagnosis

```{r}
cytomegalovirus %>% 
  dplyr::group_by(diagnosis) %>% 
  dplyr::summarise(mean(time.to.transplant, na.rm = TRUE),
                   n())
```

## Bonus material

### Advanced filter exercises

Only keep data for the following diagnoses - acute myeloid leukemia, non-Hodgkin lymphoma, myelodysplastic syndrome. Try to do this without writing multiple filter statements 

```{r}
cytomegalovirus %>% 
  dplyr::filter(diagnosis %in% c("acute myeloid leukemia", "non-Hodgkin lymphoma", "myelodysplastic syndrome"))
```

Now only keep data for the diagnoses that are NOT in the above list

```{r}
cytomegalovirus %>% 
  dplyr::filter(!(diagnosis %in% c("acute myeloid leukemia", "non-Hodgkin lymphoma", "myelodysplastic syndrome")))
```

### Selection helpers

Select all variables that **do not** have `.` in their names?

```{r}
cytomegalovirus %>% 
  dplyr::select(!contains("."))
```

Select all variables whose names end with the word "dose"

```{r}
cytomegalovirus %>% 
  dplyr::select(ends_with("dose"))
```

Select all numeric columns

```{r}
cytomegalovirus %>% 
  dplyr::select(where(is.numeric))
```

Move the `cmv` column to the start of the data frame

```{r}
# first way
cytomegalovirus %>% 
  dplyr::select(cmv, everything())

# second way
cytomegalovirus %>% 
  dplyr::relocate(cmv)
```

### Working with multiple columns

Let's work out the maximum value for each numeric variable, for each diagnosis

```{r}
cytomegalovirus %>%
  dplyr::group_by(diagnosis) %>%
  dplyr::summarise(across(where(is.numeric), ~max(.x, na.rm = TRUE))) 

# This code runs, but you get a few warnings. What do you think they mean?
```

To learn more

```{r}
vignette("colwise")
```

### Renaming variables

Change `age` to `AGE`

```{r}
cytomegalovirus %>% 
  dplyr::rename(AGE = age)
```

Replace all occurrences of "." with "-" in the relevant variable names 

```{r}
cytomegalovirus %>% 
  dplyr::rename_with(~ str_replace_all(.x, "\\.", "-") , contains("."))
```