---
title: 'Tidybiology: Class 5 Report Exercise'
output: html_document
params:
  diagnosis: "acute myeloid leukemia"
subtitle: 'My disease of interest is `r params$diagnosis`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(medicaldata)

data("cytomegalovirus")

cytomegalovirus <- cytomegalovirus %>% 
                      mutate(cmv_status = if_else(cmv == 0, "no", "yes"),
                      akir_status = if_else(aKIRs <= 4, "low", "high"),
                      akir_status = fct_relevel(akir_status, "low"))
```

## Study Background

These data were obtained from a published study that by Sobecks et al. called 
[Cytomegalovirus Reactivation After Matched Sibling Donor Reduced-Intensity Conditioning Allogeneic Hematopoietic Stem Cell Transplant Correlates With Donor Killer Immunoglobulin-like Receptor Genotype](http://www.ectrx.org/forms/ectrxcontentshow.php?year=2011&volume=9&issue=1&supplement=0&makale_no=0&spage_number=7&content_type=FULL%20TEXT)  

To learn more about the study and the data produced, run `?cytomegalovirus` in your console    

# Objective  

The goal of this study was to determine if the number of donor activating killer immunoglobulin-like receptors (aKIRs) influenced reactivation of CMV after T-cell replete, matched sibling donor reduced-intensity conditioning allogeneic hematopoietic stem cell transplant  

Donors were divided into two groups - those with either 1-4 or 5-6 alpha killer immunoglobulin-like receptors (recoded as "low" and "high" respectively by us)

CMV reactivation status is either 0 ("no") or 1 ("yes")    

```{r include=FALSE}
#####STUDENTS START HERE#####

# How many unique diagnoses are there in the `cytomegalovirus` data frame?
unique(cytomegalovirus$diagnosis)

# How many times does your disease of interest appear in this data frame?
# Get this number and save it as `num_patients`
cytomegalovirus %>% 
  count(diagnosis) %>% 
  arrange(desc(n))

num_patients <- cytomegalovirus %>% 
                  filter(diagnosis == params$diagnosis) %>% 
                  count(diagnosis) %>% 
                  arrange(desc(n)) %>% 
                  slice(n = 1) %>% 
                  pull(n)

# Create an object called `my_choice` that contains data from `cytomegalovirus` JUST FOR your disease of interest
my_choice <- cytomegalovirus %>% 
  filter(diagnosis == params$diagnosis)

# Using `my_choice`, count the number of "yes" and "no" entries for `cmv_status` 
my_choice %>% 
  count(cmv_status)

# Using `my_choice`, count the number of "yes" and "no" entries for `cmv_status` by `akir_status` group. Save this as a new object called `my_choice_counts`
my_choice_counts <- my_choice %>% 
                      group_by(akir_status) %>% 
                      count(cmv_status)
```

## Visualizing the relationship between the number of aKIR genes and CMV reactivation
```{r echo = FALSE}
# Plot `my_choice_counts` as a bar graph
my_choice_counts %>% 
  ggplot(aes(akir_status, n, fill = cmv_status)) +
  geom_col(position = position_dodge2(preserve = "single")) +
  labs(title = "Effect of aKIR status on CMV reactivation",
       subtitle = params$diagnosis) +
  theme_bw()
```

## Analysis Summary

My disease of interest is `r params$diagnosis`. There were `r num_patients` patients with this disease in this study. The total number of patients afflicted by this disease who experienced CMV reactivation was `r sum(my_choice$cmv_status == "yes")`, versus `r sum(my_choice$cmv_status == "no")` who did not. There were `r sum(my_choice$sex)` women and `r nrow(my_choice) - sum(my_choice$sex)` men with this disease, compared to `r sum(cytomegalovirus$sex)` women and `r nrow(cytomegalovirus) - sum(cytomegalovirus$sex)` men overall        
