---
title: "Advanced EDA Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary dataset - `happy_full`
```{r}
#install.packages("remotes")
#remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
library(tidybiology)

data(happy_full)
```

## Please answer the following questions

Choose all the rows between and including 50 and 100 of `happy_full`

```{r}
happy_full %>% 
  slice(n = 50:100)
```

Choose the 3 countries from each region that have the highest `generosity` scores of those regions

```{r}
happy_full %>% 
  group_by(region) %>% 
  slice_max(generosity, n = 3) %>% 
  select(country_name, generosity)
```

Select only those columns from `happy_full` that contain the word "score" in their names

```{r}
happy_full %>% 
  select(contains("score"))
```

Move the `generosity` column so that it appears after `region`

```{r}
happy_full %>% 
  relocate(generosity, .after = region)
```

Select those columns from `happy_full` that contain **more than one** underscore 

Hint: Use the stringr cheatsheet if you find this question tricky

```{r}
happy_full %>% 
  select(matches("_[^_]+_"))
```

Compute the mean `ladder_score` for each `region`. Arrange the output so that "Southeast Asia" appears first

Hint: Use `fct_relevel()`

```{r}
happy_full %>% 
  mutate(region = fct_relevel(region, "Southeast Asia")) %>% 
  group_by(region) %>% 
  summarise(mean_ladder = mean(ladder_score))
```

Plot the `generosity` scores for each region as boxplots in **ascending** order of median value

```{r}
happy_full %>% 
  ggplot(aes(fct_reorder(region, generosity), generosity)) +
  geom_boxplot()
```

Compute the means of **all** the numeric variables separately for each `region` 

```{r}
happy_full %>% 
  group_by(region) %>% 
  summarise(across(where(is.numeric), mean)) 

# bonus
happy_full %>% 
  group_by(region) %>% 
  summarise(across(where(is.numeric), mean, .names = "mean_{.col}")) 
```

Divide each numeric variable by the sum of all the numeric variables for each row. Multiply all these values by 100. Save the output data frame as `happy_100` and confirm that the row sums of the original numeric variables in this data frame all equal 100

Hint: First compute the row sums for all the numeric variables each country

```{r}
happy_100 <- happy_full %>% 
              rowwise() %>% 
              mutate(total = sum(c_across(where(is.numeric)))) %>% 
              ungroup() %>% 
              mutate(across(where(is.numeric), ~ ./total * 100)) 

# confirm that all row sums = 100
happy_100 %>% 
  select(-total) %>% 
  rowwise() %>% 
  summarise(sum(c_across(where(is.numeric))))
```

