---
title: "Advanced Filter Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary dataset - `happy_full`
```{r}
#install.packages("remotes")
#remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
library(tidybiology)

data(happy_full)
```

## Please answer the following questions

Choose countries that are NOT in Europe i.e. countries that have a `region` entry that is NOT equal to either "Western Europe" or "Central and Eastern Europe"

```{r}
happy_full %>% 
  filter(!region %in% c("Western Europe", "Central and Eastern Europe")) 

# or
happy_full %>% 
  filter(!str_detect(region, "Europe")) 
```

Choose countries that do NOT have an above average `healthy_life_expectancy` or above average `generosity` score

```{r}
happy_full %>% 
  filter(!(healthy_life_expectancy > mean(healthy_life_expectancy)|
         generosity > mean(generosity))) 
```

Choose countries that are either in Europe or have a `ladder_score` above 7, and have a `healthy_life_expectancy` above 70

```{r}
happy_full %>% 
  filter(region %in% c("Western Europe", "Central and Eastern Europe")|ladder_score > 7,
         healthy_life_expectancy > 70) 
```

Choose countries that have a `generosity` score that is between the median and mean `generosity` scores (inclusive) across all countries

```{r}
mean(happy_full$generosity)
median(happy_full$generosity)

happy_full %>% 
  filter(generosity >= median(generosity),
         generosity <= mean(generosity))
# or
happy_full %>% 
  filter(between(generosity, median(generosity), mean(generosity)))
```

Choose countries whose `ladder_score` is equal to the mean `ladder_score` +/- the standard error of `ladder_score` across all countries

Standard error formula - sd/sqrt(n)

```{r}
rows <- nrow(happy_full)

happy_full %>% 
  filter(near(ladder_score, mean(ladder_score), tol = sd(ladder_score)/sqrt(rows))) 
```

