---
title: "Dplyr Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary datasets - `happy_full` and `happy_select`
```{r}
install.packages("remotes")
remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
library(tidybiology)

data(happy_full)
data(happy_select)
```

## Please answer the following questions

How would you select just the columns `region` and `ladder_score` from the `happy_full` data frame?

```{r}
happy_full %>% 
  select(region, ladder_score)
```

How would you everything between (and including) `social_support` and `generosity`?

```{r}
happy_full %>% 
  select(social_support:generosity)
```

Select all variables that **do not** have underscores in their names?

```{r}
happy_full %>% 
  select(!contains("_"))
```

Filter the `happy_select` data frame so that it only contains information for countries in East Asia

```{r}
happy_select %>% 
  filter(region == "East Asia")
```

Now use `filter()` to only keep data for the following countries - Algeria, Belgium, India, Tunisia, and Uganda. Try to do this without writing multiple filter statements 

```{r}
happy_select %>% 
  filter(country_name %in% c("Algeria", "Belgium", "India", "Tunisia", "Uganda"))
```

Filter out information for countries that have a **below average** ladder score 

```{r}
happy_select %>% 
  filter(ladder_score > mean(ladder_score))
```

Create a new variable called `normalized_ladder_score` that contains the ladder scores for each county in `happy_full` divided by the ladder score in dystopia  

As a bonus, re-order the resulting data frame in descending order of `normalized_ladder_score` 

```{r}
happy_select %>% 
  mutate(normalized_ladder_score = ladder_score/ladder_score_in_dystopia) %>% 
  arrange(desc(normalized_ladder_score))
```

Let's now find out which are the happiest regions in the world. We'll do this by working out the average `ladder_score` of all the countries in each region 

```{r}
happy_select %>% 
  group_by(region) %>% 
  summarise(mean(ladder_score))
```

Finally, let's work out the maximum value for each numeric variable, for each region 

```{r}
happy_select %>%
  group_by(region) %>%
  summarise(across(where(is.numeric), max)) %>% View()
```


