---
title: "Ggplot2 Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
install.packages("ggExtra")
library(tidyverse)
library(ggExtra)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary datasets -  `happy_full` and `happy_select`

```{r}
install.packages("remotes")
remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
library(tidybiology)

data(happy_full)
data(happy_select)
```

## Please answer the following questions

Let's begin by getting a sense of the overall distribution of `ladder_score` in the `happy_select` data frame 

```{r}
happy_select %>% 
  ggplot(aes(ladder_score)) +
  geom_histogram()

# or
happy_select %>% 
  ggplot(aes(ladder_score)) +
  geom_density()
```

What does the `ladder_score` distribution look like for each region? 

```{r}
happy_select %>% 
  ggplot(aes(ladder_score, fill = region)) +
  geom_density()
```

Make a facet plot to visualize the distributions of `ladder_score` for each region

```{r}
happy_select %>% 
  ggplot(aes(ladder_score)) +
  geom_density() +
  facet_wrap(vars(region))
```

Make a scatter plot of `logged_gdp_per_capita` versus `ladder_score` 

```{r}
happy_select %>% 
  ggplot(aes(logged_gdp_per_capita, ladder_score)) +
  geom_point()
```

Add an appropriate trend line to the previous plot

```{r}
happy_select %>% 
  ggplot(aes(ladder_score, logged_gdp_per_capita)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

Add a title, subtitle, x-axis label, and y-axis label. Also change the overall look of the plot using your favourite `theme_*()` function 

```{r}
happy_select %>% 
  ggplot(aes(ladder_score, logged_gdp_per_capita)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Does Money == Happiness?",
       subtitle = "Plotting Ladder Score against Log GDP Per Capita",
       x = "Ladder Score",
       y = "Log GDP Per Capita") +
  theme_minimal()
```

Add histograms for `logged_gdp_per_capita` and `ladder_score` to the previous plot (to the top and to the right, respectively) 

```{r}
p1 <- happy_select %>% 
        ggplot(aes(ladder_score, logged_gdp_per_capita)) +
        geom_point() +
        geom_smooth(method = "lm", se = FALSE) +
        labs(title = "Does Money == Happiness?",
             subtitle = "Plotting Ladder Score against Log GDP Per Capita",
             x = "Ladder Score",
             y = "Log GDP Per Capita") +
        theme_minimal()

ggMarginal(
  p = p1,
  type = 'histogram'
)
```




