---
title: "Reshaping Data Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary dataset - `happy_full`
```{r}
#install.packages("remotes")
#remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
library(tidybiology)

data(happy_full)
```

## Please answer the following questions

Is `happy_full` wide or long?

```{r}
glimpse(happy_full)
```

First select the following columns - `country_name`, `region`, `ladder_score`, `healthy_life_expectancy`, and `generosity`

Next work out the mean of the three measurements for each country and call this variable `mean`

Finally, produce an output data frame that contains two columns - `country_name` and `mean`. Arrange this data frame in ascending order of mean value

```{r}
happy_full %>% 
  select(country_name, region, ladder_score, healthy_life_expectancy, generosity) %>% 
  pivot_longer(cols = !c(country_name, region),
              names_to = "measurements",
              values_to = "values") %>% 
  group_by(country_name) %>% 
  summarise(mean = mean(values)) %>% 
  arrange(mean) 
  
# versus
happy_full %>% 
  select(country_name, region, ladder_score, healthy_life_expectancy, generosity) %>% 
  rowwise() %>% 
  mutate(mean = mean(c_across(ladder_score:generosity))) %>% 
  select(country_name, mean) %>% 
  arrange(mean) 
```

First select the following columns - `country_name`, `region`, `ladder_score`, `healthy_life_expectancy`, and `generosity`

Plot a **stacked** bar graph of these three measurements for the countries Norway, India, and Thailand

Plot a **grouped** bar graph of these three measurements for the countries Norway, India, and Thailand

```{r}
# stacked
happy_long <- happy_full %>% 
                select(country_name, region, ladder_score, healthy_life_expectancy, generosity) %>% 
                pivot_longer(cols = !c(country_name, region),
                             names_to = "measurements",
                             values_to = "values") 
happy_long %>%  
  filter(country_name %in% c("Norway", "Thailand", "India")) %>% 
  ggplot(aes(country_name, values, fill = measurements)) +
  geom_col()

# grouped bar graph
happy_long %>% 
  filter(country_name %in% c("Norway", "Thailand", "India")) %>% 
  ggplot(aes(country_name, values, fill = measurements)) +
  geom_col(position = "dodge") # default is postition = "stack
```
