---
title: "Ggplot2 Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary dataset -  `mitocarta`

```{r}
install.packages("remotes")
remotes::install_github("matthewhirschey/mitocarta", force = TRUE)
library(mitocarta)
data(mitocarta)
```

## Please answer the following questions

Generate a boxplot of protein lengths for each chromosome. Hint - the relevant variables are `hg19_chromosome` and `protein_length`

```{r}

```

Log transforming protein length could help improve the previous plot. How would you do this?

```{r}

```

How would you overlay the individual points over these boxplots? 

```{r}

```

The points in the previous plot obscure the boxplots. What's one way to fix this?

```{r}

```

Boxplots are a nice way to view distributions. Another way to visualize this is by using violin plots. How would you do this?

```{r}

```

Let's make this plot nicer. How would you add the following as a plot title - "Protein length distributions across different chromosomes"? Also, remove the underscore from the y-axis label

```{r}

```

Now make the title bold

```{r}

```

Let's improve the aesthetics of this plot by using one of the pre-set themes

```{r}

```

There's a violin plot that corresponds to proteins whose chromosome language is unknown (i.e. `NA`). Remove these data points from the previous plot

```{r}

```

Make the same plot as the one above but with the axes flipped (meaning - `hg19_chromosome` is now the y-axis, and `log(protein length)` is now the x-axis)

```{r}

```

Make the same plot as above but with the chromosomes reordered i.e. chromosome 1 is at the top.

```{r}

```
