---
title: "Dplyr Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary datasets - `chromosome` and `mitocarta`
```{r}
install.packages("remotes")
remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
remotes::install_github("matthewhirschey/mitocarta", force = TRUE)
library(tidybiology)
library(mitocarta)
data(chromosome)
data(mitocarta)
```

## Please answer the following questions

How many observations and variables does `chromosome` contain?

```{r}
dim(chromosome)
# or
glimpse(chromosome)
```

What are the top five chromosomes in terms of the number of protein coding genes (look at the `protein_codinggenes` variable)? Based on your output, would you say that chromosome length is proportional to the number of protein coding genes?

```{r}
chromosome %>% 
  arrange(desc(protein_codinggenes)) %>% 
  slice(1:5)
```

Using the `protein_codinggenes` and `pseudo_genes` variables, create a new variable called `total_genes` that contains the sum of protein coding genes and pseudogenes. Save the output of this as a new object, also called `chromosome`

```{r}
chromosome <- chromosome %>% 
                mutate(total_genes = protein_codinggenes + pseudo_genes)
```

How many chromosomes have a higher than average total number of genes? 

```{r}
chromosome %>% 
  filter(total_genes > mean(total_genes)) %>% 
  nrow()
```

Take a glimpse at the `mitocarta` dataframe

```{r}
glimpse(mitocarta)
```

How many *unique* entries are there in the column `mito_domain_score`?

```{r}
mitocarta %>% 
  distinct(mito_domain_score)
```

Remove all rows that have `NA` values in the `mito_domain_score` column. Replace `mitocarta` with this new dataframe

```{r}
mitocarta <- mitocarta %>% 
               filter(mito_domain_score != "NA")
# alternatively
mitocarta <- mitocarta %>% 
               drop_na(mito_domain_score)
```

What is the average length of all the proteins in `mitocarta`?

```{r}
mitocarta %>% 
  summarise(mean_length = mean(protein_length))
```

What is the average protein length for each of the three `mito_domain_score` categories?

```{r}
mitocarta %>% 
  group_by(mito_domain_score) %>% 
  summarise(mean_length = mean(protein_length))
```

