---
title: "Ggplot2 Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary dataset -  `mitocarta`

```{r}
install.packages("remotes")
remotes::install_github("matthewhirschey/mitocarta", force = TRUE)
library(mitocarta)
data(mitocarta)
```

## Please answer the following questions

Generate a boxplot of protein lengths for each chromosome. Hint - the relevant variables are `hg19_chromosome` and `protein_length`

```{r}
ggplot(mitocarta, aes(x = hg19_chromosome, y = protein_length)) +
  geom_boxplot()
```

Log transforming protein length could help improve the previous plot. How would you do this?

```{r}
ggplot(mitocarta, aes(x = hg19_chromosome, y = log(protein_length))) +
  geom_boxplot()
```
How would you overlay the individual points over these boxplots? 
```{r}
ggplot(mitocarta, aes(x = hg19_chromosome, y = log(protein_length))) +
  geom_boxplot() +
  geom_jitter()
```
The points in the previous plot obscure the boxplots. What's one way to fix this?
```{r}
ggplot(mitocarta, aes(x = hg19_chromosome, y = log(protein_length))) +
  geom_boxplot(outlier.shape = NA) + # this gets rid of the outlier points produces by `geo_boxplot(`
  geom_jitter(alpha = 0.1)
```
Boxplots are a nice way to view distributions. Another way to visualize this is by using violin plots. How would you do this?
```{r}
ggplot(mitocarta, aes(x = hg19_chromosome, y = log(protein_length))) +
  geom_violin()
```

Let's make this plot nicer. How would you add the following as a plot title - "Protein length distributions across different chromosomes"? Also, re,pve the underscore from the y-axis label
```{r}
ggplot(mitocarta, aes(x = hg19_chromosome, y = log(protein_length))) +
  geom_violin() +
  labs(title = "Protein length distributions across different chromosomes",
       y = "log(protein length)")
```
Now make the title bold
```{r}
ggplot(mitocarta, aes(x = hg19_chromosome, y = log(protein_length))) +
  geom_violin() +
  labs(title = "Protein length distributions across different chromosomes",
       y = "log(protein length)") +
  theme(plot.title = element_text(face = "bold"))
```
Let's improve the aesthetics of this plot by using one of the pre-set themes
```{r}
ggplot(mitocarta, aes(x = hg19_chromosome, y = log(protein_length))) +
  geom_violin() +
  labs(title = "Protein length distributions across different chromosomes",
       y = "log(protein length)") +
  theme_minimal() + 
  theme(plot.title = element_text(face = "bold")) # note: this MUST come after `theme_minimal()` if you want the text to continue to be bold
```

There's a violin plot that corresponds to proteins whose chromosome language is unknown (i.e. `NA`). Remove these data points from the previous plot
```{r}
mitocarta %>% 
  filter(hg19_chromosome != "NA") %>% 
  ggplot(aes(x = hg19_chromosome, y = log(protein_length))) +
  geom_violin() +
  labs(title = "Protein length distributions across different chromosomes",
       y = "log(protein length)") +
  theme_minimal() + 
  theme(plot.title = element_text(face = "bold"))
```

