---
title: "Join Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse")
library(tidyverse)
```

## Required packages and datasets

Run the code in the code chunk below to load the necessary dataset -  `proteins` and `mitocarta`

```{r}
install.packages("remotes")
remotes::install_github("matthewhirschey/tidybiology", force = TRUE)
remotes::install_github("matthewhirschey/mitocarta", force = TRUE)
library(tidybiology)
library(mitocarta)
data(proteins)
data(mitocarta)

# keep just a few columns in `proteins` and `mitocarta`
proteins <- proteins %>% 
              select(gene_name, sequence)
mitocarta <- mitocarta %>% 
              select(symbol, mito_domain_score) %>% 
              drop_na(mito_domain_score)
```

## Please answer the following questions

Which columns in `proteins` and `mitocarta` can we use to join on?
```{r}
glimpse(proteins)
glimpse(mitocarta)
# We see that `gene_name` from `proteins` and `symbol` from `mitocarta` contain the same information, and so we use this to join on
```

Combine `proteins` and `mitocarta`, ensuring that ALL the rows from `proteins` are retained in the resulting dataframe. Save this as a new dataframe called `proteins_mito`
```{r}
proteins_mito <- proteins %>% 
                  left_join(mitocarta, by = c("gene_name" = "symbol")) 
```

How many missing values does `proteins_mito` contain in its `mito_domain_score` column?
```{r}
proteins_mito %>% 
  count(mito_domain_score)
```

Combine `proteins` and `mitocarta`, ensuring that there are no missing values in the `mito_domain_score` column of the resulting dataframe (call this new dataframe `mito_proteins`)
```{r}
mito_proteins <- proteins %>% 
                  right_join(mitocarta, by = c("gene_name" = "symbol")) 

mito_proteins %>% 
  count(mito_domain_score) # there are now no missing values in this column
```

Combine `proteins` and `mitocarta` in such a way that the resulting dataframe contains no missing values. Do not save the output but do confirm that there are no missing values in the output
```{r}
proteins %>% 
  inner_join(mitocarta, by = c("gene_name" = "symbol")) %>% 
  nrow() # 15631

proteins %>% 
  inner_join(mitocarta, by = c("gene_name" = "symbol")) %>% 
  drop_na() %>% 
  nrow() # also 15631, confirming that there are no missing values in the output
```

